---
title: "Graph Creation"
author: "Jaya Vellayan"
date: "3/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r}
library(plotly)
library(DT)
library(ggplot2)
library(tidyverse)
library(rio)


```


```{r}
data <- read.csv("EconomistData.csv")
data
```

#psuedocode
1)create updated_regions column with actual names used in graphic
2)plot x = CPI, y = HDI, color = updated_region with geom_point
  - change fill of points to white, and make the outline thicker
  - add axis names
    -bold title
    -italicize titles
3)manually change the axis to match the graphic
4)remove excess background lines on the graph
6)add r^2 line using geosmooth
  -manually add label for this, and change color to red
7)use ggrepel to add country labels
8)remove legend title




```{r}

region_lkup <- c(Americas="Americas", MENA="Middle East & North Africa", `EU W. Europe`="OECD", `East EU Cemt Asia`="Central & Eastern Europe", `Asia Pacific`="Asia and Oceania", SSA="Sub-Saharan")

data$updated_regions <- as.character(region_lkup[data$Region])

data

```


```{r}
final_remake1 <- ggplot(data = data, 
       mapping = aes(x = CPI,
                    y = HDI
                    )
                ) +
              geom_point(mapping = aes(color = updated_regions), size =3, fill="white", shape=1, stroke=2) 
final_remake1

```


```{r}
final_remake1 <- final_remake1 + labs(x = 'Corruption Perceptions Index, 2011 (10=least corrupt)',
y = 'Human Development Index, 2011 (1=best)',
title = 'Corruption and human development') + 
theme(axis.title.x = element_text(size=8)) +
theme(axis.title.y = element_text(size=8)) 


final_remake1
```
```{r}
final_remake1 <- final_remake1 + theme(plot.title = element_text(face="bold"))
final_remake1
```


```{r}
final_remake1 <- final_remake1 + scale_x_continuous(limits=c(1,10), n.breaks = 10) + scale_y_continuous(limits=c(0.2,1.0), n.breaks = 9) + theme_bw()
final_remake1
```

```{r}
final_remake1 <- final_remake1  + theme(panel.grid.major.x=element_blank(), 
                     panel.grid.minor.x=element_blank(), panel.grid.minor.y =element_blank())

final_remake1

```



```{r}
final_remake1 <- final_remake1 + theme(axis.title.x=element_text(face="italic")) + theme(axis.title.y=element_text(face="italic")) + theme(plot.title = element_text(face="bold"))
final_remake1
```

```{r}
final_remake1 <- final_remake1 + theme(legend.position='top')
final_remake1
```


```{r}
final_remake1 <- final_remake1 + geom_smooth(method = "lm", formula = y ~ log(x), color = "red", se = FALSE, span=0.5)
final_remake1
```




```{r}
#install.packages("ggrepel")
library(ggrepel)
```

```{r}
library(ggrepel)
final_remake1 <- final_remake1 + geom_text_repel(aes(label = Country), color = "black", data = subset(data, Country %in% c("Russia", "Venezuela", "Iraq", "Myanmar", "Sudan",
                   "Afghanistan", "Congo", "Greece", "Argentina", "Brazil",
                   "India", "Italy", "China", "South Africa", "Spane",
                   "Botswana", "Cape Verde", "Bhutan", "Rwanda", "France",
                   "United States", "Germany", "Britain", "Barbados", "Norway",
                   "Japan", "New Zealand", "Singapore")),check_overlap = TRUE)

final_remake1
```





```{r}

final_remake1 <- final_remake1 + scale_color_manual(values=c("#33CCFF", "#00F7FF", "#0EAEAE", "#26649A", "#EC5A25", "#98413A"))

final_remake1
```


```{r}
w <-0.56
final_remake1 <- final_remake1 + annotate(geom = 'text', 
                                        x = 9, y = 0.25, 
                                        label=glue::glue("R^2 == {l}",l=w),
                                        size=4,
                                        color='red',
                                        parse=T)
final_remake1
```


```{r}
final_remake1 <- final_remake1 + theme(legend.title=element_blank())

final_remake1
```


```{r}
save.image(file ="final_remake1.RData")

```
